---
title: "Getting Started with cjdiag"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Getting Started with cjdiag}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 9.6,
  fig.height = 5.4,
  dpi = 300,
  out.width = "100%"
)
```

## 1. What cjdiag does

Standard conjoint analysis tools estimate Average Marginal Component
Effects (AMCEs) --- the causal effect of changing a single attribute
level. AMCEs tell you *what* respondents prefer, but not *how* they
decide: which attribute levels they actually attend to, which ones they
ignore, and in what order they process information.

**cjdiag** fills that gap. It works at the level of individual
**attribute levels** --- not aggregated attributes --- because the
specific level (e.g., "no plans to look for work", not just "Job Plans"
as a whole) is what triggers respondent decisions.

cjdiag is complementary to `cjoint`/`cregg` (which estimate AMCEs).
The intended workflow is: run those for AMCEs first, then run cjdiag
to diagnose how respondents actually made those choices.

## 2. Quick start

```{r setup}
library(cjdiag)
data(immig)

f <- Chosen_Immigrant ~ Gender + Education + LanguageSkills +
  CountryofOrigin + Job + JobExperience + JobPlans +
  ReasonforApplication + PriorEntry

rf <- cj_fit(f, data = immig, method = "forest")
rf
```

```{r quick-plot, fig.height = 7.2}
plot(rf, type = "rank", top_n = 20)
```

## 3. Choosing a method

| Estimand | `method =` | Question | Output | Behavioural assumption | When to use |
|----------|-----------|----------|--------|-----------------------|-------------|
| **Level importance** | `"forest"` | Which attribute levels matter most? | MDA, root-split rate per level | None --- non-parametric | Default. Always fit this first. |
| **Decision structure** | `"tree"` | How do respondents structure their decisions? | Hierarchical CART splits | Lexicographic / sequential | When you suspect a gatekeeper. |
| **Level attendance** | `"crt"` | Which levels survive a strict signal-vs-noise test? | Lambda-survival, attended/ignored | Sparsity (most levels are noise) | When you want a hard attendance test. |
| **Decision order** | `"nmm"` | In what order do levels settle choices? | Decisiveness ranking, cumulative % | Sequential elimination (EBA) | When you care about the decision *order*. |
| **Individual attendance** | `"marginal_r2"` | Which attributes did each respondent actually use? | Per-respondent R² matrix | Per-respondent simple-regression fit | When you want individual-level heterogeneity. |

Each method also has its own task-oriented vignette:
[forest](forest.html), [tree](tree.html), [nmm](nmm.html),
[marginal_r2](marginal_r2.html), [crt](crt.html).

## 4. `cj_fit()` --- full reference

`cj_fit()` is the single entry point. It dispatches on `method` and
returns an S3 object inheriting from `cjdiag_fit`.

### Signature

```r
cj_fit(formula, data, method = c("forest", "tree", "crt", "nmm",
                                  "marginal_r2"),
       resolution = c("levels", "attributes"),
       ntree = 500L, cp = 0.005,
       lambda_grid = c(1, 2, 3, 5, 7, 10, 15, 20, 25, 30,
                       40, 50, 75, 100, 150, 200, 300, 400, 500),
       n_folds = 5L, n_perm = 20L, tol = 1e-3,
       resp_id = NULL, n_boot = 0L,
       seed = 42L, ...)
```

### Required arguments

- **`formula`** --- a formula `outcome ~ attr1 + attr2 + ...`. The
  outcome must be binary (0/1, logical, or a 2-level factor);
  predictors are treated as categorical (converted to factors
  internally).
- **`data`** --- a data frame in long format, i.e. one row per profile.

### Method choice

- **`method`** --- one of `"forest"` (default), `"tree"`, `"crt"`,
  `"nmm"`, `"marginal_r2"`. See §3.

### Resolution

- **`resolution`** --- `"levels"` (default) or `"attributes"`.
  - `"levels"` dummy-codes every attribute level and treats each as a
    separate predictor. The recommended default --- importance is a
    property of *specific levels*, not aggregated attributes.
  - `"attributes"` passes the original factor columns to the model.
    Available for `forest` and `tree` only. CRT, NMM and `marginal_r2`
    require `"levels"` and will error otherwise.

### Method-specific arguments

| Argument | Methods | Default | Purpose |
|---|---|---|---|
| `ntree` | `forest` | `500L` | Number of trees in the random forest. |
| `cp` | `tree` | `0.005` | rpart complexity parameter. Smaller = deeper trees. |
| `lambda_grid` | `crt` | 19 values from 1 to 500 | L1 regularization path for hierNet. **Must extend well past 50** --- typical signal-vs-noise separation occurs in the 100--500 range. |
| `n_folds` | `crt` | `5L` | Cross-validation folds. |
| `n_perm` | `crt` | `20L` | Permutation rounds for permutation importance. |
| `tol` | `crt` | `1e-3` | hierNet convergence tolerance. |
| `resp_id` | `nmm`, `marginal_r2` | `NULL` | **Required.** Name of the respondent ID column. |
| `n_boot` | `nmm` | `0L` | Bootstrap iterations for NMM CIs. `0` disables. |

### Reproducibility

- **`seed`** --- random seed; default `42L`. Set explicitly for any run
  you want to reproduce.

### Pass-through

- **`...`** --- additional arguments forwarded to
  `randomForest::randomForest()`, `rpart::rpart()`, or
  `hierNet::hierNet()` depending on `method`.

### Errors and validation

`cj_fit()` errors with a structured `cli` message in these cases:

- `data` is not a data frame, or has zero rows.
- `method = "crt"` or `"nmm"` is combined with `resolution = "attributes"`.
- `method = "nmm"` or `"marginal_r2"` is called without `resp_id`.
- The formula's outcome is not binary.
- An attribute named in the formula is missing from `data`.

### Return value

An S3 object. The exact subclass depends on `method`:
`cjdiag_forest`, `cjdiag_tree`, `cjdiag_crt`, `cjdiag_nmm`,
`cjdiag_marginal_r2`. All inherit from `cjdiag_fit` and support
`print()`, `plot()`, `summary()`, and `importance()`.

The object is a list. Common slots: `model` (the underlying fit),
`method`, `resolution`, `results` (the importance tibble), `formula`,
`outcome`, `attributes`, `n_obs`, `n_levels`, `attr_map`, `seed`.
Method-specific slots (`oob_error`, `ntree` on forest;
`root_split`, `depth`, `n_terminal`, `cp` on tree;
`optimal_lambda`, `lambda_1se`, `accuracy` on crt; `n_boot` on nmm;
`r2_matrix` on marginal_r2) round it out.

## 5. `importance()` --- extracting the results table

`importance()` is the S3 generic that returns a `cjdiag_importance`
object --- a tidy tibble of per-level importance metrics plus a few
summary slots. Methods are dispatched on the fit's class.

```{r importance}
imp <- importance(rf)
imp
```

Coerce to a data frame for downstream work:

```{r importance-df}
head(as.data.frame(imp))
```

Per-method slots on the returned object:

- **forest**: `results`, `root_dist`, `oob_error`, `ntree`.
- **tree**: `results`, `root_split`, `depth`.
- **crt**: `results`, `optimal_lambda`, `lambda_1se`, `n_attended`,
  `accuracy`.
- **nmm**: `results`, `n_boot`.
- **marginal_r2**: `results`, `n_resp`, `r2_matrix`.

`importance()` makes no copies of the underlying model --- it is a
thin projection of `fit$results`. For analyses that need the model
itself, use `fit$model`.

## 6. `plot()` --- every plot type

All plot methods return `ggplot` objects (except the tree, which
renders via `rpart.plot::rpart.plot()` and returns invisibly). All
accept the customization parameters in §8.

### 6.1 `plot.cjdiag_forest()`

```r
plot(forest_obj, type = c("importance", "combined", "rank"),
     top_n = NULL, ...)
```

- `type = "importance"` (default) --- vertical bar chart of MDA per
  level, sorted descending.
- `type = "combined"` --- dual-axis plot showing MDA bars + root-split
  rate as overlaid points/line. Useful for comparing the two
  importance signals at a glance.
- `type = "rank"` --- line plot of MDA against rank, with rotated
  x-tick labels (attribute black, level dark grey). The default for
  the README and the plot type used in §11.

### 6.2 `plot.cjdiag_tree()`

```r
plot(tree_obj, ...)
```

Renders the decision tree via `rpart.plot::rpart.plot()`. Splits are
relabelled `attribute: level` (e.g. `JobPlans: no plans to look for work`)
via a custom `split.fun`. Pass any `rpart.plot` argument through `...`
to override defaults (`type = 4`, `extra = 101`, `under = TRUE`,
`fallen.leaves = TRUE`, `box.palette = "BuOr"`, `cex = 0.9`).

### 6.3 `plot.cjdiag_crt()`

```r
plot(crt_obj, type = c("robustness", "survival", "rank", "mda", "cv"),
     top_n = NULL, ...)
```

- `type = "robustness"` (default) --- bar chart of `max_lambda` per
  level (the largest λ where the level survives).
- `type = "survival"` --- λ-survival curve: which levels keep a
  non-zero coefficient as λ rises.
- `type = "rank"` --- connected-dot plot of `max_lambda` per level,
  ordered from most to least attended. Same aesthetic as
  `plot(forest, type = "rank")` but the y-axis is the survival
  statistic instead of MDA.
- `type = "mda"` --- permutation importance from the CRT fit.
- `type = "cv"` --- cross-validated accuracy across the lambda grid.

### 6.4 `plot.cjdiag_nmm()`

```r
plot(nmm_obj, top_n = NULL, ...)
```

Cumulative-explanation curve only --- the single supported NMM plot
since 0.2.1. Earlier `type = "decisiveness"` and `type = "sample"`
plots were removed (the columns remain on the fit object).

### 6.5 `plot.cjdiag_importance()` (generic)

```r
plot(imp_obj, type = c("mda", "root", "combined", "cumulative",
                       "cumulative_pct"),
     top_n = NULL, ...)
```

Plots from a `cjdiag_importance` object directly. Valid types depend
on the original method:

| Original method | Valid types |
|---|---|
| forest | `mda`, `root`, `combined`, `cumulative`, `cumulative_pct` |
| crt | `mda` |
| tree | `mda` |
| nmm | `mda`, `cumulative` |
| marginal_r2 | `mda` |

For most users, `plot()` directly on the `cj_fit()` output is simpler
than going through `importance()`.

### 6.6 Common plot arguments

| Argument | Type | Purpose |
|---|---|---|
| `top_n` | integer or `NULL` | Number of levels to display. Default `25L`; `NULL` = all levels. |
| `base_size` | numeric | Base font size. Falls back to global option, then `12`. |
| `colors` | named char vector | Override specific palette colors (e.g. `c(primary = "#000")`). |
| `palette` | char | `"default"`, `"colorblind"` (Okabe-Ito), or `"grey"`. |
| `theme` | `theme()` object | Replace the default theme entirely. |
| `label_wrap` | integer | Character width for label wrapping. Default `35L`. |
| `attribute.names` | named char vector | Per-call attribute renames (e.g. `c(JobPlans = "Plans for Employment")`). |
| `level.names` | named list | Per-call level renames. See §8.2. |
| `...` | --- | Forwarded to the underlying geom or `rpart.plot`. |

## 7. `print()` and `summary()`

Every fit and importance object has a `print()` method tuned to the
method:

- **`print.cjdiag_forest()`** --- header (resolution, ntree, OOB error,
  observations, attributes/levels) and the top `n` rows of the
  results table, including `mda`, `root_pct`, `class_0`, `class_1`.
- **`print.cjdiag_tree()`** --- header (cp, root split, depth, terminal
  nodes) and the top `n` rows by importance.
- **`print.cjdiag_crt()`** --- header (optimal λ, λ-1se, accuracy,
  attended count) and the top `n` rows by `max_lambda`.
- **`print.cjdiag_nmm()`** --- header (total pairs, decisive picks)
  and the top `n` levels by decisiveness.
- **`print.cjdiag_marginal_r2()`** --- per-attribute mean/median R²
  and share of respondents with R² = 0.
- **`print.cjdiag_importance()`** --- same shape, dispatched on the
  importance object.

All print methods accept an `n` argument:

```{r print-n}
print(rf, n = 5)
```

Without `n`, `print()` falls back to `get_cjdiag_options("print_n")`
(default `10L`).

`summary()` methods exist for `cjdiag_forest`, `cjdiag_tree`,
`cjdiag_crt`, `cjdiag_nmm`, and `cjdiag_importance`. They return a
list of summary statistics rather than printing --- useful when you
want the numbers without the formatted output.

## 8. Customization

### 8.1 Theme

`theme_cjdiag()` is the default ggplot2 theme used internally:

```r
theme_cjdiag(base_size = 12, base_family = "",
             grid_y = FALSE, grid_x = TRUE)
```

Built on `theme_minimal()` with no minor gridlines, optional major
gridlines on x or y axes, and the legend at the top.

### 8.2 Per-call overrides

Every plot method accepts the customization arguments from §6.6.
Examples:

```{r custom-plot, fig.height = 7.2}
plot(rf,
     palette = "colorblind",
     attribute.names = c(LanguageSkills = "English Proficiency",
                         JobPlans = "Plans for Employment"),
     top_n = 20)
```

Renaming levels uses a list keyed by attribute:

```{r custom-levels, eval = FALSE}
plot(rf,
     level.names = list(
       JobPlans = c("no plans to look for work" = "no plans",
                    "contract with employer" = "has contract")
     ))
```

### 8.3 Global options

For the same overrides across many plots in a session or script, set
them once globally:

```{r global-theme, eval = FALSE}
set_cjdiag_theme(base_size = 14,
                 palette = "colorblind",
                 print_n = 5)

set_cjdiag_labels(
  attribute.names = c(LanguageSkills = "English Proficiency"),
  level.names = list(
    JobPlans = c("no plans to look for work" = "no plans")
  )
)
```

`set_cjdiag_theme()` arguments:

| Argument | Default | Purpose |
|---|---|---|
| `base_size` | `12` | Font size. |
| `palette` | `"default"` | One of `"default"`, `"colorblind"`, `"grey"`. |
| `font_family` | `""` | Base font family. |
| `label_wrap` | `35L` | Label-wrap character width. |
| `theme` | `NULL` | Full ggplot2 theme override. |
| `print_n` | `10L` | Default `n` for `print()` methods. |

`set_cjdiag_labels()` arguments:

| Argument | Purpose |
|---|---|
| `attribute.names` | Named char vector mapping original to display. |
| `level.names` | Named list of per-attribute level remaps. |
| `reset` | If `TRUE`, clears the dictionary. |

Both setters return the previous options invisibly, so you can save
and restore around a block:

```{r save-restore, eval = FALSE}
old <- set_cjdiag_theme(palette = "grey")
on.exit(do.call(set_cjdiag_theme, old))
```

Inspect current options:

```{r get-options}
get_cjdiag_options("palette")
```

`get_cjdiag_options()` with no argument returns the full options list.

### 8.4 Palettes

`cjdiag_palette()` returns the named character vector for a palette:

```{r palette}
cjdiag_palette("colorblind")
```

The `n` argument is reserved for future use; currently each palette
returns three colors (`primary`, `secondary`, `tertiary`).

The three palettes:

- `"default"` --- blue / red / light-grey.
- `"colorblind"` --- Okabe-Ito blue / vermillion / grey.
- `"grey"` --- dark / mid / light grey, for grayscale print.

### 8.5 Resolution priority

For every plot argument, the priority is:

```
explicit function arg  >  global option (set_cjdiag_*)  >  hardcoded default
```

This lets you set sensible defaults once and override per-plot when
needed.

## 9. `augment_profile_order()` --- utility for CRT

CRT requires the *profile order* assumption (the choice probability
must be invariant to which side a profile is shown on). Most
real-world conjoint data don't satisfy this exactly; the standard fix
is to flip left/right and invert the outcome, doubling the dataset.

```r
augment_profile_order(data, outcome,
                      left = c("left_attr1", "left_attr2", ...),
                      right = c("right_attr1", "right_attr2", ...))
```

- `data` --- a data frame.
- `outcome` --- name of the binary outcome column (0/1).
- `left`, `right` --- vectors of attribute column names, same length.
  Pairs must align: `left[i]` is swapped with `right[i]`.

Returns a data frame with `2 × nrow(data)` rows: the original data
followed by the swapped/inverted copy. Apply this **before**
`cj_fit(..., method = "crt")` if your data isn't already symmetric.

This is not needed for `forest`, `tree`, `nmm`, or `marginal_r2`.

## 10. Datasets

`data(immig)` --- the bundled Hainmueller and Hopkins (2015)
immigration conjoint. 13,960 profile evaluations across 1,396
respondents, 9 attributes:

```{r immig-summary}
str(immig)
```

The outcome `Chosen_Immigrant` is binary (0 = rejected, 1 = chosen).
`CaseID` is the respondent identifier (use as `resp_id` for `nmm` and
`marginal_r2`).

## 11. Worked workflow

A representative end-to-end session combining several methods:

```{r workflow, fig.height = 6}
# 1. Importance ranking - always start here.
rf <- cj_fit(f, data = immig, method = "forest")
top_levels <- importance(rf)$results
head(top_levels[, c("rank", "attribute", "level", "mda", "root_pct")], 10)

# 2. Decision structure - does a single attribute gate the rest?
tr <- cj_fit(f, data = immig, method = "tree")
tr$root_split   # the gatekeeper, if any

# 3. Decision order - sequential elimination view.
nmm <- cj_fit(f, data = immig, method = "nmm",
              resp_id = "CaseID", n_boot = 0)

# 4. Per-respondent attendance - heterogeneity diagnostic.
mr2 <- cj_fit(f, data = immig, method = "marginal_r2",
              resp_id = "CaseID")
```

```{r workflow-plot, fig.height = 7.2}
plot(rf, type = "rank", top_n = 15)
```

Each method asks a different question. They are complementary, not
competing.

## 12. Reproducibility checklist

- Set `seed` explicitly (default is `42L`, but make this explicit in
  any saved analysis script).
- For `forest`, `ntree = 500L` is plenty for stable rankings;
  `ntree = 1000L` is more conservative.
- For `crt`, the lambda grid must extend well beyond 50 --- the
  default goes to 500 and most signal-vs-noise separation happens
  between 100 and 500. Truncating at λ = 50 underestimates how many
  levels survive penalization. `n_perm = 20L` is the published
  default; raise to `100--500` if you intend to publish individual
  lambda-survival numbers.
- For `nmm`, `n_boot = 0` returns point estimates only; set to
  `200` (or `1000`) to attach bootstrap CIs to the decisiveness
  ranks.
- All `print()` and `plot()` arguments are deterministic given the
  fit; only `cj_fit()` itself uses randomness.

## 13. Citation

```{r citation}
citation("cjdiag")
```
