---
title: "CRT / HierNet: Which levels survive a strict signal-vs-noise test?"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CRT / HierNet: Which levels survive a strict signal-vs-noise test?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 9.6,
  fig.height = 5.4,
  dpi = 300,
  out.width = "100%"
)
```

## When to use

Use `method = "crt"` when you want a hard *attendance test*: which
attribute levels keep a non-zero coefficient under increasing L1
penalty? Levels that vanish quickly under penalization are noise or
redundant; levels that survive the heaviest penalties genuinely drive
choices.

CRT requires the [hierNet](https://cran.r-project.org/package=hierNet)
package. It is the most computationally expensive method in cjdiag.

```{r setup}
library(cjdiag)
data(immig)

f <- Chosen_Immigrant ~ Gender + Education + LanguageSkills +
  CountryofOrigin + Job + JobExperience + JobPlans +
  ReasonforApplication + PriorEntry
```

## Fit

```{r crt-fit, eval = requireNamespace("hierNet", quietly = TRUE)}
crt <- cj_fit(f, data = immig, method = "crt",
              lambda_grid = c(5, 10, 20, 50, 100, 150, 200, 300, 500),
              n_folds = 3, n_perm = 5)
crt
```

`max_lambda` is the largest λ at which the level still has a non-zero
coefficient. `attended` is a boolean derived from `max_lambda > 0`.

Most signal-vs-noise separation happens between λ = 100 and λ = 500.
Stopping the grid at λ = 50 (or lower) is a common mistake that
makes nearly every level look "attended": the penalty isn't yet
strong enough to eliminate the noise. Always extend the grid past
λ = 200; the package default goes to 500.

## Plot lambda survival

Two complementary views of the survival statistic.

`type = "survival"` shows each level's coefficient path across the
lambda grid, so you can see at what penalty a level drops out.

```{r crt-survival, eval = requireNamespace("hierNet", quietly = TRUE), fig.height = 7.2}
plot(crt, type = "survival")
```

`type = "rank"` collapses each path to its `max_lambda` and plots the
levels in descending order — a one-number attendance ranking in the
same connected-dot style as `plot(forest, type = "rank")`.

```{r crt-rank, eval = requireNamespace("hierNet", quietly = TRUE), fig.height = 5.4}
plot(crt, type = "rank")
```

## Related

- [Marginal R-squared](marginal_r2.html) for a per-respondent
  attendance signal.
- [Random Forest](forest.html) for a non-parametric importance ranking
  (no penalization).
