---
title: "Random Forest: Which attribute levels matter most?"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Random Forest: Which attribute levels matter most?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 9.6,
  fig.height = 5.4,
  dpi = 300,
  out.width = "100%"
)
```

## When to use

Always fit `method = "forest"` first. It is non-parametric, makes no
behavioural assumptions, and gives you the importance hierarchy of
individual attribute *levels*.

```{r setup}
library(cjdiag)
data(immig)

f <- Chosen_Immigrant ~ Gender + Education + LanguageSkills +
  CountryofOrigin + Job + JobExperience + JobPlans +
  ReasonforApplication + PriorEntry
```

## Fit

```{r forest-fit}
rf <- cj_fit(f, data = immig, method = "forest")
rf
```

## Interpret the results table

```{r forest-results}
rf$results
```

Key columns:

- **mda** --- Mean Decrease in Accuracy. How much the forest's predictive
  accuracy drops when this level's values are shuffled. Higher = more
  important. Levels with `mda` near zero are effectively ignored.
- **root_pct** --- % of trees in the forest where this level is the
  *first* split. The **gatekeeper signal**.
- **class_0** / **class_1** --- class-conditional importance. `class_0`
  large = level mostly drives *rejection* (deal-breaker); `class_1`
  large = level mostly drives *selection* (attractor).

## Plot

```{r forest-plot, fig.height = 7.2}
plot(rf, type = "rank", top_n = 20)
```

## Related

- [Decision Tree](tree.html) for the hierarchical structure of these importances.
- [Nested Marginal Means](nmm.html) for the decision *order*.
