---
title: "Marginal R-squared: Which attributes did each respondent actually use?"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Marginal R-squared: Which attributes did each respondent actually use?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 9.6,
  fig.height = 5.4,
  dpi = 300,
  out.width = "100%"
)
```

## When to use

Use `method = "marginal_r2"` (Jenke, Bansak, Hainmueller and Hangartner
2021) when you care about *individual-level* heterogeneity in attribute
attendance --- which respondents used which attributes. The other
methods aggregate across respondents; this one does not.

For each respondent, the method regresses their choices on each
attribute *alone* and reports the R². A respondent with R² near zero
for an attribute likely ignored it entirely.

```{r setup}
library(cjdiag)
data(immig)

f <- Chosen_Immigrant ~ Gender + Education + LanguageSkills +
  CountryofOrigin + Job + JobExperience + JobPlans +
  ReasonforApplication + PriorEntry
```

## Fit

```{r mr2-fit}
mr2 <- cj_fit(f, data = immig, method = "marginal_r2", resp_id = "CaseID")
mr2
```

The print output gives the mean / median R² per attribute and the
share of respondents with R² = 0 (the per-attribute non-attendance
rate).

## Per-respondent matrix

The full per-respondent × per-attribute R² matrix is on the fit object:

```{r mr2-matrix}
str(mr2$individual_r2)
```

You can summarise this however you like --- e.g. count the number of
attributes each respondent used (R² above some threshold) to study
heterogeneity in attendance breadth.

## Related

- [CRT / HierNet](crt.html) for an aggregate, regularization-based
  attendance test.
- [Random Forest](forest.html) for population-level importance.
