---
title: "Nested Marginal Means: In what order do levels settle choices?"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Nested Marginal Means: In what order do levels settle choices?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 9.6,
  fig.height = 5.4,
  dpi = 300,
  out.width = "100%"
)
```

## When to use

Use `method = "nmm"` (Dill, Howlett & Mueller-Crepon 2024) when you
care about the *order* in which attribute levels settle choices, not
just their static importance. NMM is the closest match to elimination-
by-aspects (Tversky 1972) at the level of attribute levels.

The procedure works sequentially:

1. Identify the level whose marginal mean deviates most from 50/50 ---
   the most *decisive* level.
2. Remove choice tasks where that level cannot discriminate (because
   both profiles share it).
3. Repeat on the reduced sample.

The cumulative plot shows how quickly the top levels account for the
total decisiveness.

```{r setup}
library(cjdiag)
data(immig)

f <- Chosen_Immigrant ~ Gender + Education + LanguageSkills +
  CountryofOrigin + Job + JobExperience + JobPlans +
  ReasonforApplication + PriorEntry
```

## Fit

```{r nmm-fit}
nmm <- cj_fit(f, data = immig, method = "nmm",
              resp_id = "CaseID", n_boot = 0)
nmm
```

## Plot the cumulative explanation curve

```{r nmm-plot, fig.height = 7.2}
plot(nmm, top_n = 20)
```

A steep early curve = a strong decision-order hierarchy: a few
top-ranked levels settle most of the choices. A flat curve = compensatory
processing, where many levels each contribute a little.

## Related

- [Decision Tree](tree.html) for an alternative hierarchical
  representation that uses CART splits instead of marginal means.
- [Random Forest](forest.html) for static level importance without an
  ordering assumption.
