---
title: "Decision Tree: How do respondents structure their decisions?"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Decision Tree: How do respondents structure their decisions?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 9.6,
  fig.height = 5.4,
  dpi = 300,
  out.width = "100%"
)
```

## When to use

Use `method = "tree"` when you suspect respondents apply a hierarchical
elimination rule --- a lexicographic *gatekeeper* attribute they check
first, then a smaller cue conditional on it. The root split identifies
the gatekeeper. The tree uses only a subset of available levels,
consistent with respondents processing a few key cues rather than all
information at once.

```{r setup}
library(cjdiag)
data(immig)

f <- Chosen_Immigrant ~ Gender + Education + LanguageSkills +
  CountryofOrigin + Job + JobExperience + JobPlans +
  ReasonforApplication + PriorEntry
```

## Fit

```{r tree-fit}
tr <- cj_fit(f, data = immig, method = "tree")
tr
```

## Plot

```{r tree-plot}
plot(tr)
```

Splits are labelled `Attribute: level` (e.g. `JobPlans: no plans to
look for work`) so the diagnostic is readable without per-dataset
customisation.

## Related

- [Random Forest](forest.html) gives the population-level importance
  hierarchy that the tree's splits draw from.
- [Nested Marginal Means](nmm.html) for an alternative
  sequential-elimination view that does not require fitting a tree.
